\chapter{History}

\section{timeline}
\begin{itemize}
  \item 1930 -- first electronic component on a Vehicle
  \item TODO
  \item 1980 -- OBD is introduced
\end{itemize}

Today up to 2500 singals are exchanged through up to 150 ECUs, over at least 5 types of busses. Ford F-150 has 150M lines of codes onboard.

\section{ADAS vs AD}
ADAS -- Advanced Driver Assistance System -- are technologies for driver's safety or convenience. The responsibility os leaved to the user of the system.

AD -- Autonomous Driving -- takes the responsibility away from the driver and puts it in the system.

Notice the difference between logitudinal (accelerator) and lateral (path) control.

See DARPA challenge.

ODD - Operational Design Domain - defines the scenarios where the vehicle must be functional.

\section{AD Classification}

\subsection{SAE Autonomous Driving Levels}
\begin{itemize}
  \item 0. No automation
  \item 1. Driver automation: adaptive cruise control, line keeping
  \item 2. Partial assistance: both cruse control and line keeping
  \item 3. TODO
  \item 4. TODO
  \item 5. Full autonomous: unlimited ODD
\end{itemize}

\subsection{Types of AD}
Many types of AD approaches exist: geometrical approach (classical one), end-to-end approach based on deep learning, LLM- and multimodal- based models.

A classical AD pipeline is comprised of several sequential stages:
\begin{enumerate}
  \item Sensing: sense the environment (sensors and V2V signals)
  \item Perception: broke up in environmental perception and localization
    \begin{enumerate}
      \item Environmentlal perception: ranges from easy solutions (aka detect and consider still) up to 
        \begin{enumerate}
          \item object detection
          \item object classification
        \end{enumerate}
      \item localization
    \end{enumerate}
  \item Planning: plans a path; this task is iterated multiple times at different Levels
    \begin{enumerate}
      \item mission/global planning
      \item behavioral planning
      \item motion/local planning
    \end{enumerate}
  \item Control: tries to follow the path with a control algo
    \begin{itemize}
      \item Longitudinal control
      \item Lateral control
    \end{itemize}
\end{enumerate}

Different sensors are used for creating ADAS and AD applications: LiDAR, radar, cameras, IMUs and GPS receivers.

Other kinds of approaches to the AD problem are being explored. Machine learning based approaches are primarly divided into full end-to-end and hybrid end-to-end.
\begin{itemize}
  \item Full end-to-end directly maps each sensor to each actutor and lets the hidden layer of the NN learn complex patterns without decomposing the main problem into subproblems
  \item Hybrid end-to-end mixes classical stages with NN-based stages. Eg.: the localization might still be provided by a classical pipeline, while some of the other stages might be implemeted through a NN 
\end{itemize}

\section{Players}
\begin{enumerate}
  \item Waymo
  \item Beidu
  \item Cruse
  \item Zoox
  \item Pony.AI
\end{enumerate}